---
title: "The vibe of musical and social reward: Listening to beat-based music as a surrogate for socioemotional support during the Covid-19 pandemic across Europe"
date: "`r format(Sys.time(), '%d %B, %Y')`"
author:
- "Lena Esther Ptasczynski [Charité - Universitätsmedizin Berlin, lena-esther.ptasczynski@charite.de]"
- "Patrick Blättermann [Institute of Sound and Vibration Engineering, Hochschule Düsseldorf, patrick.blaettermann@hs-duesseldorf.de]"
- "Fabian Greb [independent researcher, fabian.greb@gmx.de]"
- "Philipp Sterzer [University Psychiatric Clinics Basel, philipp.sterzer@upk.ch]"
- "Jochen Steffens [Institute of Sound and Vibration Engineering, Hochschule Düsseldorf, jochen.steffens@hs-duesseldorf.de]"
output:
  github_document:
    toc: true
  html_document:
    toc: true
---

# Description

This script runs mixed-effects mediation analyses for danceability, valence, and energy, with the Stringency Index as predictor, temperature as covariate, and multiple mediators. Additionally, it performs a pre/post analysis comparing the period before versus after the onset of the Covid-19 pandemic, and computes country-level descriptive statistics.

```{=html}
<style>
.main-container {
  max-width: 95% !important;
  margin-left: 20px !important;
  margin-right: 20px !important;
}

.container-fluid {
  max-width: 95% !important;
}

body {
  max-width: 100%;
}
</style>
```



```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)

is_html_output <- knitr::is_html_output(excludes = c("markdown", "gfm"))
```


# Load libraries

```{r}
# Install libraries if not installed already
# and load libraries
for (package in c("dplyr", "tibble", "performance", "lme4", "lmerTest", "mediation", "purrr", "knitr", "DT")) {
  if (!require(package, character.only = TRUE, quietly = TRUE)) {
    install.packages(package)
    library(package, character.only = TRUE)
  }
}
```

# Helper functions

```{r}

build_formula <- function(response, predictors) {
  as.formula(
    paste(response, "~", paste(predictors, collapse = " + "))
  )
}

format_effect <- function(est, p, sig, digits = 3) {
  p_txt <- ifelse(
    is.na(p),
    "NA",
    format.pval(p, digits = digits, eps = .001)
  )

  out <- sprintf(
    paste0("%.", digits, "f%s (p = %s)"),
    est, sig, p_txt
  )

  out[is.na(est)] <- NA_character_
  out
}

p_to_stars <- function(p) {
  if (is.na(p)) return("")
  if (p < .001) return("***")
  if (p < .01)  return("**")
  if (p < .05)  return("*")
  if (p < .1)   return(".")
  ""
}

extract_coef_info <- function(model, term) {
  coef_tab <- coef(summary(model))

  if (!term %in% rownames(coef_tab)) {
    return(tibble::tibble(
      estimate = NA_real_,
      p_value = NA_real_,
      sig = NA_character_
    ))
  }

  p_col <- grep("^Pr\\(", colnames(coef_tab), value = TRUE)

  p_val <- if (length(p_col) == 1) coef_tab[term, p_col] else NA_real_

  tibble::tibble(
    estimate = unname(coef_tab[term, "Estimate"]),
    p_value = unname(p_val),
    sig = p_to_stars(unname(p_val))
  )
}

extract_model_effect_info <- function(model,
                                      term = "Stringency_index",
                                      conf_level = 0.95,
                                      ci_method = "Wald") {

  coef_tab <- coef(summary(model))

  if (!term %in% rownames(coef_tab)) {
    return(tibble::tibble(
      estimate = NA_real_,
      p_value = NA_real_,
      ci_lower = NA_real_,
      ci_upper = NA_real_,
      sig = NA_character_
    ))
  }

  p_col <- grep("^Pr\\(", colnames(coef_tab), value = TRUE)
  p_val <- if (length(p_col) == 1) coef_tab[term, p_col] else NA_real_

  ci_vals <- tryCatch(
    {
      ci_obj <- suppressMessages(
        confint(model, parm = term, level = conf_level, method = ci_method)
      )
      as.numeric(ci_obj[1, ])
    },
    error = function(e) c(NA_real_, NA_real_)
  )

  tibble::tibble(
    estimate = unname(coef_tab[term, "Estimate"]),
    p_value = unname(p_val),
    ci_lower = ci_vals[1],
    ci_upper = ci_vals[2],
    sig = p_to_stars(unname(p_val))
  )
}

extract_med_info <- function(med_object,
                             total_model = NULL,
                             total_term = "Stringency_index") {
  s <- summary(med_object)

  indirect_est <- if ("d.avg" %in% names(s)) s$d.avg else if ("d0" %in% names(s)) s$d0 else NA_real_
  indirect_p   <- if ("d.avg.p" %in% names(s)) s$d.avg.p else if ("d0.p" %in% names(s)) s$d0.p else NA_real_

  direct_est   <- if ("z.avg" %in% names(s)) s$z.avg else if ("z0" %in% names(s)) s$z0 else NA_real_
  direct_p     <- if ("z.avg.p" %in% names(s)) s$z.avg.p else if ("z0.p" %in% names(s)) s$z0.p else NA_real_

  if (!is.null(total_model)) {
    total_info <- extract_model_effect_info(total_model, term = total_term)
    total_est <- total_info$estimate
    total_p   <- total_info$p_value
  } else {
    total_est <- if ("tau.coef" %in% names(s)) s$tau.coef else NA_real_
    total_p   <- if ("tau.p" %in% names(s)) s$tau.p else NA_real_
  }

  tibble::tibble(
    indirect_effect = indirect_est,
    indirect_p = indirect_p,
    indirect_sig = p_to_stars(indirect_p),

    direct_effect = direct_est,
    direct_p = direct_p,
    direct_sig = p_to_stars(direct_p),

    total_effect = total_est,
    total_p = total_p,
    total_sig = p_to_stars(total_p)
  )
}

make_summary_row <- function(result, analysis_id, section_title, outcome, mediator,
                             treatment = "Stringency_index") {

  med_info <- extract_med_info(
    med_object = result$med,
    total_model = result$m3_lmtest,
    total_term = treatment
  )

  a_info <- extract_coef_info(result$m2_lmtest, treatment)
  b_info <- extract_coef_info(result$m1_lmtest, mediator)

  tibble::tibble(
    analysis_id = analysis_id,
    section_title = section_title,
    outcome = outcome,
    mediator = mediator,

    indirect_effect = med_info$indirect_effect,
    indirect_p = med_info$indirect_p,
    indirect_sig = med_info$indirect_sig,

    direct_effect = med_info$direct_effect,
    direct_p = med_info$direct_p,
    direct_sig = med_info$direct_sig,

    total_effect = med_info$total_effect,
    total_p = med_info$total_p,
    total_sig = med_info$total_sig,

    a_path = a_info$estimate,
    a_p = a_info$p_value,
    a_sig = a_info$sig,

    b_path = b_info$estimate,
    b_p = b_info$p_value,
    b_sig = b_info$sig
  )
}

fit_model_set <- function(data, outcome, mediator,
                          treatment = "Stringency_index",
                          group = "Country",
                          covariate = "Temp_mean") {

  # same formulas as in your original code
  f_m1 <- build_formula(outcome, c(treatment, mediator, paste0("(1|", group, ")"), covariate))
  f_m2 <- build_formula(mediator, c(treatment, paste0("(1|", group, ")")))
  f_m3 <- build_formula(outcome, c(treatment, paste0("(1|", group, ")"), covariate))
  f_m4 <- build_formula(outcome, c(mediator, paste0("(1|", group, ")"), covariate))

  # lme4 models
  m1 <- lme4::lmer(f_m1, data = data)
  m2 <- lme4::lmer(f_m2, data = data)
  m3 <- lme4::lmer(f_m3, data = data)
  m4 <- lme4::lmer(f_m4, data = data)

  # mediation model
  set.seed(21)
  med <- mediation::mediate(
    model.m = m2,
    model.y = m1,
    treat = treatment,
    mediator = mediator
  )

  # lmerTest models for significance tests
  m1_lmtest <- lmerTest::lmer(f_m1, data = data)
  m2_lmtest <- lmerTest::lmer(f_m2, data = data)
  m3_lmtest <- lmerTest::lmer(f_m3, data = data)
  m4_lmtest <- lmerTest::lmer(f_m4, data = data)

  list(
    m1 = m1, m2 = m2, m3 = m3, m4 = m4,
    med = med,
    m1_lmtest = m1_lmtest, m2_lmtest = m2_lmtest,
    m3_lmtest = m3_lmtest, m4_lmtest = m4_lmtest
  )
}

get_first_existing <- function(x, candidates) {
  for (nm in candidates) {
    if (!is.null(x[[nm]])) return(x[[nm]])
  }
  NULL
}

extract_med_effect_from_object <- function(med_object,
                                           effect = c("indirect", "direct", "total"),
                                           conf_level = 0.95) {
  effect <- match.arg(effect)

  alpha <- 1 - conf_level
  probs <- c(alpha / 2, 1 - alpha / 2)

  if (effect == "indirect") {
    effect_label <- "Indirect effect (ACME)"
    est <- get_first_existing(med_object, c("d.avg", "d0"))
    p   <- get_first_existing(med_object, c("d.avg.p", "d0.p"))
    sims <- get_first_existing(med_object, c("d.avg.sims", "d0.sims"))
  } else if (effect == "direct") {
    effect_label <- "Direct effect (ADE)"
    est <- get_first_existing(med_object, c("z.avg", "z0"))
    p   <- get_first_existing(med_object, c("z.avg.p", "z0.p"))
    sims <- get_first_existing(med_object, c("z.avg.sims", "z0.sims"))
  } else {
    effect_label <- "Total effect"
    est <- get_first_existing(med_object, c("tau.coef"))
    p   <- get_first_existing(med_object, c("tau.p"))
    sims <- get_first_existing(med_object, c("tau.sims"))
  }

  ci_lower <- NA_real_
  ci_upper <- NA_real_

  if (!is.null(sims)) {
    ci_vals <- stats::quantile(as.numeric(sims), probs = probs, na.rm = TRUE)
    ci_lower <- unname(ci_vals[1])
    ci_upper <- unname(ci_vals[2])
  }

  tibble::tibble(
    Effect = effect_label,
    Estimate = as.numeric(est)[1],
    ci_lower = ci_lower,
    ci_upper = ci_upper,
    p_value = as.numeric(p)[1],
    Significance = p_to_stars(as.numeric(p)[1])
  )
}

format_ci <- function(lower, upper, digits = 3) {
  out <- sprintf(
    paste0("[%.", digits, "f, %.", digits, "f]"),
    lower, upper
  )
  out[is.na(lower) | is.na(upper)] <- NA_character_
  out
}

make_med_ci_table <- function(result,
                              digits = 3,
                              total_term = "Stringency_index",
                              total_ci_method = "Wald") {

  indirect <- extract_med_effect_from_object(result$med, "indirect")

  direct <- extract_med_effect_from_object(result$med, "direct")

  total_model_info <- extract_model_effect_info(
    model = result$m3_lmtest,
    term = total_term,
    ci_method = total_ci_method
  ) %>%
    dplyr::transmute(
      Effect = "Total effect",
      Estimate = estimate,
      ci_lower = ci_lower,
      ci_upper = ci_upper,
      p_value = p_value,
      Significance = sig
    )

  dplyr::bind_rows(
    indirect,
    direct,
    total_model_info
  ) %>%
    dplyr::mutate(
      `95% CI` = format_ci(ci_lower, ci_upper, digits = digits),
      Estimate = round(Estimate, digits),
      `p-value` = round(p_value, digits)
    ) %>%
    dplyr::select(
      Effect,
      Estimate,
      `95% CI`,
      `p-value`,
      Significance
    )
}
```

# Load dataframes

```{r}
data_all_europe_st <- readRDS("data/data_all_europe_st.RDS")

data_all_stress_europe <- readRDS("data/data_all_stress_europe.RDS")

data_all_stress_europe_st <- readRDS("data/data_all_stress_europe_st.RDS")

```

# Calculate means and SEM

```{r}
means <- data_all_stress_europe %>%
            group_by(Country) %>%
            summarise(
              Mean_danceability = mean(Danceability_mean, na.rm=TRUE),
              Mean_stringency_index = mean(Stringency_index, na.rm=TRUE),
              Mean_sps_sum = mean(SPS_SUM_mean, na.rm=TRUE),
              Mean_sps_si = mean(SPS_SI_mean, na.rm=TRUE),
              Mean_sps_row = mean(SPS_ROW_mean, na.rm=TRUE),
              Mean_sps_a = mean(SPS_A_mean, na.rm=TRUE),
              Mean_sps_sora = mean(SPS_SORA_mean, na.rm=TRUE),
              Mean_sps_g = mean(SPS_G_mean, na.rm=TRUE),
              SEM_danceability = sd(Danceability_mean, na.rm=TRUE)/sqrt(sum(!is.na(Danceability_mean))),
              SEM_stringency_index = sd(Stringency_index, na.rm=TRUE)/sqrt(sum(!is.na(Stringency_index))),
              SEM_sps_sum = sd(SPS_SUM_mean, na.rm=TRUE)/sqrt(sum(!is.na(SPS_SUM_mean))),
              SEM_sps_si = sd(SPS_SI_mean, na.rm=TRUE)/sqrt(sum(!is.na(SPS_SI_mean))),
              SEM_sps_row = sd(SPS_ROW_mean, na.rm=TRUE)/sqrt(sum(!is.na(SPS_ROW_mean))),
              SEM_sps_a = sd(SPS_A_mean, na.rm=TRUE)/sqrt(sum(!is.na(SPS_A_mean))),
              SEM_sps_sora = sd(SPS_SORA_mean, na.rm=TRUE)/sqrt(sum(!is.na(SPS_SORA_mean))),
              SEM_sps_g = sd(SPS_G_mean, na.rm=TRUE)/sqrt(sum(!is.na(SPS_G_mean)))
  )

means
```

# Pre/post analysis

```{r}
model_period_dance_lmtest = lmerTest::lmer(Danceability_mean ~ relevel(Period, ref="pre")+Temp_mean+(1|Country), data = data_all_europe_st)
model_period_dance = lme4::lmer(Danceability_mean ~ relevel(Period, ref="pre")+Temp_mean+(1|Country), data = data_all_europe_st)

set.seed(21)
model_period_dance_CI = confint(model_period_dance, method="boot", nsim=1000)

summary(model_period_dance_lmtest)
performance::r2(model_period_dance_lmtest)
performance::icc(model_period_dance_lmtest)

model_period_dance_CI

##Check regression Assumptions (visual)
plot(model_period_dance_lmtest)

qqnorm(resid(model_period_dance_lmtest))
hist(resid(model_period_dance_lmtest))
```

# Create analysis plan for mediation

```{r}
analysis_plan <- tribble(
  ~analysis_id,                ~section_title,                                                  ~outcome,              ~mediator,                     ~med_summary_output, ~med_plot,
  "string_sps_dance",          "Stringency -> SPS SUM Score -> Dance",                          "Danceability_mean",   "SPS_SUM_mean",                NA_character_,       FALSE,
  "string_si_dance",           "Stringency -> SPS Social Integration -> Dance",                 "Danceability_mean",   "SPS_SI_mean",                 NA_character_,       FALSE,
  "string_row_dance",          "Stringency -> SPS Reassurance of worth score -> Dance",         "Danceability_mean",   "SPS_ROW_mean",                NA_character_,       FALSE,
  "string_a_dance",            "Stringency -> SPS Attachment -> Dance",                         "Danceability_mean",   "SPS_A_mean",                  NA_character_,       FALSE,
  "string_sora_dance",         "Stringency -> SPS Sense of reliable alliance -> Dance",         "Danceability_mean",   "SPS_SORA_mean",               NA_character_,       FALSE,
  "string_g_dance",            "Stringency -> SPS Guidance -> Dance",                           "Danceability_mean",   "SPS_G_mean",                  NA_character_,       FALSE,
  "string_slon_dance",         "Stringency -> Scale SLON SUM -> Dance",                         "Danceability_mean",   "Scale_SLON_SUM_mean",         NA_character_,       FALSE,
  "string_pss_dance",          "Stringency -> Scale PSS10 UCLA SUM_mean -> Dance",              "Danceability_mean",   "Scale_PSS10_UCLA_SUM_mean",   NA_character_,       FALSE,

  "string_sps_valence",        "Stringency -> SPS SUM Score -> Valence",                        "Valence_mean",        "SPS_SUM_mean",                NA_character_,       FALSE,
  "string_si_valence",         "Stringency -> SPS Social Integration -> Valence",               "Valence_mean",        "SPS_SI_mean",                 NA_character_,       FALSE,
  "string_row_valence",        "Stringency -> SPS Reassurance of worth score -> Valence",       "Valence_mean",        "SPS_ROW_mean",                NA_character_,       FALSE,
  "string_a_valence",          "Stringency -> SPS Attachment -> Valence",                       "Valence_mean",        "SPS_A_mean",                  NA_character_,       FALSE,
  "string_sora_valence",       "Stringency -> SPS Sense of reliable alliance -> Valence",       "Valence_mean",        "SPS_SORA_mean",               NA_character_,       FALSE,
  "string_g_valence",          "Stringency -> SPS Guidance -> Valence",                         "Valence_mean",        "SPS_G_mean",                  NA_character_,       FALSE,

  "string_sps_energy",         "Stringency -> SPS SUM Score -> Energy",                         "Energy_mean",         "SPS_SUM_mean",                NA_character_,       FALSE,
  "string_si_energy",          "Stringency -> SPS Social Integration -> Energy",                "Energy_mean",         "SPS_SI_mean",                 NA_character_,       FALSE,
  "string_row_energy",         "Stringency -> SPS Reassurance of worth score -> Energy",        "Energy_mean",         "SPS_ROW_mean",                NA_character_,       FALSE,
  "string_a_energy",           "Stringency -> SPS Attachment -> Energy",                        "Energy_mean",         "SPS_A_mean",                  NA_character_,       FALSE,
  "string_sora_energy",        "Stringency -> SPS Sense of reliable alliance -> Energy",        "Energy_mean",         "SPS_SORA_mean",               NA_character_,       FALSE,
  "string_g_energy",           "Stringency -> SPS Guidance -> Energy",                          "Energy_mean",         "SPS_G_mean",                  NA_character_,       FALSE
)

kable(
  analysis_plan %>%
    dplyr::select(section_title, outcome, mediator),
  col.names = c("Analysis", "Outcome", "Mediator")
)
```

# Fit mediation models

```{r}
all_results_list <- vector("list", nrow(analysis_plan))

pb <- txtProgressBar(min = 0, max = nrow(analysis_plan), style = 3)

for (i in seq_len(nrow(analysis_plan))) {
  all_results_list[[i]] <- fit_model_set(
    data = data_all_stress_europe_st,
    outcome = analysis_plan$outcome[[i]],
    mediator = analysis_plan$mediator[[i]]
  )

  setTxtProgressBar(pb, i)
}

close(pb)

all_results <- analysis_plan %>%
  dplyr::mutate(result = all_results_list)

mediation_summary_table <- purrr::pmap_dfr(
  list(
    result = all_results$result,
    analysis_id = all_results$analysis_id,
    section_title = all_results$section_title,
    outcome = all_results$outcome,
    mediator = all_results$mediator
  ),
  make_summary_row
) %>%
  dplyr::mutate(
    dplyr::across(
      c(
        indirect_effect, indirect_p,
        direct_effect, direct_p,
        total_effect, total_p,
        a_path, a_p,
        b_path, b_p
      ),
      ~ round(.x, 4)
    )
  )
```

# Mediation results

```{r}
mediation_summary_table_display <- mediation_summary_table %>%
  dplyr::transmute(
    Analysis = section_title,
    Outcome = outcome,
    Mediator = mediator,
    `Indirect effect (ACME)` = format_effect(indirect_effect, indirect_p, indirect_sig),
    `Direct effect (ADE)`   = format_effect(direct_effect, direct_p, direct_sig),
    `Total effect`          = format_effect(total_effect, total_p, total_sig),
    `a path: Stringency → Mediator` = format_effect(a_path, a_p, a_sig),
    `b path: Mediator → Outcome`    = format_effect(b_path, b_p, b_sig)
  )

if (is_html_output) {
  DT::datatable(
    mediation_summary_table_display,
    rownames = FALSE,
    filter = "top",
    options = list(
      paging = FALSE,
      scrollX = TRUE,
      scrollY = "600px",
      scrollCollapse = TRUE,
      autoWidth = TRUE
    ),
    caption = "Summary of mediation analyses"
  )
} else {
  knitr::kable(
    mediation_summary_table_display,
    caption = "Summary of mediation analyses"
  )
}
```

# Mediation results (CI)

```{r}

all_mediation_ci_table <- purrr::pmap_dfr(
  list(
    result = all_results$result,
    section_title = all_results$section_title,
    outcome = all_results$outcome,
    mediator = all_results$mediator
  ),
  function(result, section_title, outcome, mediator) {
    make_med_ci_table(result) %>%
      dplyr::mutate(
        Analysis = section_title,
        Outcome = outcome,
        Mediator = mediator,
        .before = 1
      )
  }
)

if (is_html_output) {
  DT::datatable(
    all_mediation_ci_table,
    width = "100%",
    rownames = FALSE,
    filter = "top",
    class = "display nowrap",
    options = list(
      dom = "lfrtip",
      paging = TRUE,
      lengthChange = TRUE,
      pageLength = 20,
      lengthMenu = list(
        c(10, 20, 50, 100, -1),
        c("10", "20", "50", "100", "All")
      ),
      scrollX = TRUE,
      scrollY = "500px",
      scrollCollapse = TRUE,
      autoWidth = TRUE
    ),
    caption = "Confidence intervals for all mediation models"
  )
} else {
  knitr::kable(
    all_mediation_ci_table,
    caption = "Confidence intervals for all mediation models"
  )
}

```

# Session info

```{r}
sI=sessionInfo()
print(sI, RNG = FALSE, locale = FALSE)
```